Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-21 paper/uist_2016/paper.tex
minor edit with some added quotes

\subsubsection{Rewinds vs. Static Photos} 
Our study reveals that participants have higher semantic memory of locations when presented with Rewinds than with static street-view photos ($\chi^2(1, N=356) = 96.987$, $p <.01$). When presented with a street-view photo generated from the participant's location data, participants were able to recognize the depicted location 49.7\% of the time. Recognition rate of locations reached 96.6\% when participants were shown the accompanying Rewind. As discussed above, this can be expected because cinematic media like Rewinds provide more context and retrieval cues. Three participants (P4, P6, and P10) who were unable to recall a location from a static photo actually recognized the location after seeing the accompanying Rewind. They expressed their contentment after watching the Rewind, proclaiming ``I recognize the photo now!'' When we evaluate Rewind memorability based on episodic memory rather than semantic memory, the results show that participants recalled the exact trip depicted by the Rewind 53.7\% of the time.
%TODO make sure episodic memory is discussed in discussion section--visit frequency diminishes episodic memory

@@

\subsubsection{Visit Frequency}
% only affect location rememberance
% trip rememberance $\chi^2(1,126) = 0.066$, p = 0.3987 (> 0.05)
% most frequently visited - 72%, least frequently visited - 36%, $\chi^2(1,128) = 14.458$, p < 0.01
Location memorability and trip memorability are affected differently by frequency of visits (Table~\ref{table:betweensubjects}). Location memorability is highest for places that are most frequently visited by the participant, while trip memorability is highest for places that are least frequently visited. The more frequently a place was visited, the more likely it was that the participant remembered the location when it was shown to them at the study. In the most frequently visited places, participants remembered the location 72\% of the time, while only remembering the location 36\% of the time in least frequently visited places, with a significance at $\chi^2(1,128) = 14.458$, p < 0.01. This effect does not occur for trip memorability. In the most frequently visited places, trips are remembered 51\% of the time. Instead of seeing trip memorability decrease for least frequently visited places, this factor increases to 56\%.

Our study examined episodic processing in metric-specific groups, as well as the performance overall. Our study shows that comparing to static street-view images, recollection of event-driven memories using Rewinds is significantly improved for trips taken that were 100 - 1,000 km (62 - 620 miles) away from home by 50\%, $\chi^2(1, N=20) = 3.215$, $p <.05$; and for trips that were least frequently taken by about 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$.

@@

% Our findings show that GPS usage significantly affect the rememberance of locations $\chi^2(1,179) = 8.734$, p < 0.01, but does not impose a siginificant effect on trip rememberance. (GPS users remember locations 33% of the time whereas non-GPS users remember locations 58% of the time)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% heather - come back to this if time allows
% because memories are simply not encoded so there's no way to retrieve it in the first place?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our findings show that GPS usage significantly affect participants' remembrance of locations $\chi^2(1,179) = 8.734$, p = 0.0016 ($< 0.01$), but does not impose significant effects on trip remembrance $\chi^2(1,177) = 0.713$, p = 0.1992 ($> 0.05$).  When asked whether they remember the location in a static photo, non-regular GPS users gave a positive answer 58\% of the time whereas regular GPS users gave a positive answer 33\% of the time. Although the difference between non-regular and regular GPS users' trip remembrance is not statistically significance, the between-subject percentage values still lean towards the observation that it is perhaps easier for non-GPS users to recall episodic memories from external memory cues, given that the total trip remembrance rate for non-GPS users is 48\% and that of GPS users is 56\%. Since regular GPS users are frequently directed step-by-step on how to navigate through their environment, regular GPS users may not need to pay as much attention to their surroundings, leading to potentially diminished locational awareness which could be the reason why regular GPS users had a lower remembrance rate for their ambience, particularly locations. \\

@@

\subsubsection{Travel Companions}
% trip accompanies have siginificant effects on both location and trip rememberance. 
% location rememberance: travel alone -  100%, not alone - 55%, $\chi^2(1,94) = 31.812$, p < 0.01
% trip rememberance: travel alone - 62%, not alone - 90%, $\chi^2(1,133) = 9.016$, p < 0.01
Our study shows that both location and trip remembrances are influenced by the presence of travel companions, yet in a opposite way. For location rememberance, participants who believed that they traveled alone reported 100\% successful recalls of locations compare to 55\% rate reported by participants who believed that they traveled in groups, with $\chi^2(1,94) = 31.812$, $p < 0.00001$. On the other hand, trip details tend to be more easily remembered during accompanied visits. Participants who believed that they traveled alone were able to remember the exact trip 62\% of the time from watching Rewinds, yet those who believed that they traveled with companions could remember the exact trip 90\% of the time, with $\chi^2(1,133) = 9.016$, p = 0.0013 $(< 0.01)$.

@@

%TODO where to put this????  maybe split discussion into three sections on mem/believ/desir
% Jeff thinks that the discussion is already a bit long. It's fine in here, but maybe shorten it. Don't add more sections
\subsection{Episodic Memory}
It could possibly be inferred that participants are likely to benefit more from Rewinds than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip rememberance when presented with static images as opposed to what they were asked after watchiing Rewinds, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip; since if participants fail to form any cognitive association with the locations depicted in the images, it is very unlikely that they would be able to associate target itineraries with the locations shown. Thus, it is perhaps safe to presume that location rememberance with images bounds above the maximum likelihood of trip rememberance with the same images presented to the participants, and we used location rememberance from image recognition as an upper-limit gauge against trip rememberance from Rewinds thereafter.

One plausible explanation of the improvement may be inherent to the encoding of episodic memory in the first place. As suggested by the levels-of-processing theory(ref), encoding efficacy is heavily dependent on the levels to which stimuli are being actively or passively processed. Not surprisingly, superficial mental engagement would produce weaker, and less enduring cognitive representation which would demand more contextually accurate and abundant cues for memory triggery, and vice versa. Oftentimes, the cognitive representation resulted from shallow processing may not even be retrievable at all as it is simply not available for retrieval.  As previously described, both of the trips revealed in our results, those that were taken farthest away from home and those that were least frequently taken, may not be special enough to engage deep memory encoding nor often enough to reinforce incidental remembrance. This means that it would be more difficult for participants to recall these trips if the provided cues are not effective enough. Since Rewinds provide richer and more contextually meaningful cues than static street-side images, it is reasonable that Rewinds significantly outperforms plain images as suggested by our statistical analysis. While being asked about location and trip rememberance on one of his least frequently taken routes, P7 was not able to identify the location from the static image at first; yet he recognized the trip from Rewinds as ``one street (in the video) helps me (him) remember". 
% The Effects of Divided Attention on Encoding and Retrieval Processes in Human Memory 
% http://www2.psych.ubc.ca/~pgraf/Psy583Readings/Craik%20et%20al%201996.pdf
% Craik et al. 1996

The extent to which attention is dedicated while taking a trip may be another possible explanation for the observed improvement using Rewinds. This is perhaps particularly applicable to explain why participants who were self-identified regular GPS users tend to benefit more from Rewinds with episodic memory recalling. Laboratory work done by Craik et al. 1996 (ref) has demonstrated that the encoding of memory is more effective under full-attention conditions than divided-attention conditions.  This means that regular GPS users, who are often directed step-by-step on how to navigate through their environments, may be less likely to engage their attention while taking their routes, leading to less available cognitive representation for retrieval. P3, a self-identified regular GPS user who had 30\% recalling rate using static images and 80\% trip remembrance rate using Rewinds said ``I really don't pay attention to my surroundings when I'm driving. (I am) Focused (focused) on GPS and road"  Under this circumstance, it is almost certain that Rewinds would provide a more successful access to episodic memory as it provides more spatial-temporal cues for memory retrieval, not only by adjusting image attributes to reflect timely and weather changes, but also by mimicing the motions of the users\' when they travel down a particular path.

It may appear to be puzzling that participants seem to recall less about specifics of their trips from Rewinds compared to static imageries when they traveled alone, with a significant decrease around 40\%, $\chi^2(1, N=168) = 33.831$, $p <.01$; when the trips were most frequently taken, with a significant decrease of 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$; and when the trips taken were 10 - 100 meters (33 - 328 feet) away from home, with a mildly significant decrease of 50\%, $\chi^2(1, N=16) = 2.25$, $p <.1$. One possible reason for this would be it is hard for participants to tie a repeatedly taken trip with a specific instance in time, and trips taken near home could also be the ones that are most frequently taken. However, since our study was using location rememberance from images as an upper-limit estimation about participants' trip rememberance from the images, it is also likely that the actual trip rememberance from images could be lower.

>>>>

% h - deleted participants # for consistencies - P4, P6 and P10 for future reference
\subsubsection{Rewinds vs. Static Photos} 
Our study reveals that participants have higher semantic memory of locations when presented with Rewinds than with static street-view photos ($\chi^2(1, N=356) = 96.987$, $p <.01$). When presented with a street-view photo generated from the participant's location data, participants were able to recognize the depicted location 49.7\% of the time. Recognition rate of locations reached 96.6\% when participants were shown the accompanying Rewind. As discussed above, this can be expected because cinematic media like Rewinds provide more context and retrieval cues. Three participants who were unable to recall a location from a static photo actually recognized the location after seeing the accompanying Rewind. They expressed their contentment after watching the Rewind, proclaiming ``I recognize the photo now!'' When we evaluate Rewind memorability based on episodic memory rather than semantic memory, the results show that participants recalled the exact trip depicted by the Rewind 53.7\% of the time.
%TODO make sure episodic memory is discussed in discussion section--visit frequency diminishes episodic memory

@@

\subsubsection{Visit Frequency}
% only affect location rememberance
% trip rememberance $\chi^2(1,126) = 0.066$, p = 0.3987 (> 0.05)
% most frequently visited - 72%, least frequently visited - 36%, $\chi^2(1,128) = 14.458$, p < 0.01
Location memorability and trip memorability are affected differently by frequency of visits (Table~\ref{table:betweensubjects}). Location memorability is highest for places that are most frequently visited by the participant, while trip memorability is highest for places that are least frequently visited. The more frequently a place was visited, the more likely it was that the participant remembered the location when it was shown to them at the study. In the most frequently visited places, participants remembered the location 72\% of the time, while only remembering the location 36\% of the time in least frequently visited places, with a significance at $\chi^2(1,128) = 14.458$, $p < 0.01$. This effect does not occur for trip memorability. In the most frequently visited places, trips are remembered 51\% of the time. Instead of seeing trip memorability decrease for least frequently visited places, this factor increases to 56\%. Our study also shows that comparing to static street-view images, recollection of trip memories using Rewinds is significantly improved for least frequently taken trips by about 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$.

% Our study examined episodic processing in metric-specific groups, as well as the performance overall. Our study shows that comparing to static street-view images, recollection of event-driven memories using Rewinds is significantly improved for trips taken that were 100 - 1,000 km (62 - 620 miles) away from home by 50\%, $\chi^2(1, N=20) = 3.215$, $p <.05$; and for trips that were least frequently taken by about 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$.

@@

% Our findings show that GPS usage significantly affect the rememberance of locations $\chi^2(1,179) = 8.734$, p < 0.01, but does not impose a siginificant effect on trip rememberance. (GPS users remember locations 33% of the time whereas non-GPS users remember locations 58% of the time)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% heather - come back to this if time allows
% because memories are simply not encoded so there's no way to retrieve it in the first place?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Our findings show that GPS usage significantly affect participants' remembrance of locations $\chi^2(1,179) = 8.734$, p = 0.0016 ($< 0.01$), but does not impose significant effects on trip remembrance $\chi^2(1,177) = 0.713$, p = 0.1992 ($> 0.05$).  When asked whether they remember the location in a static photo, non-regular GPS users gave a positive answer 58\% of the time whereas regular GPS users gave a positive answer 33\% of the time. Although the difference between non-regular and regular GPS users' trip remembrance is not statistically significance, the between-subject percentage values still lean towards the observation that it is perhaps easier for non-GPS users to recall episodic memories from external memory cues, given that the total trip remembrance rate for non-GPS users is 48\% and that of GPS users is 56\%. Since regular GPS users are frequently directed step-by-step on how to navigate through their environment, they may not need to pay as much attention to their surroundings, leading to potentially diminished locational awareness which could be the reason why regular GPS users had a lower remembrance rate for their ambience, particularly locations. One non-regular GPS user mentioned that ``I identify by landmarks'', yet this is something that regular GPS users may pay less attention to. \\

@@

\subsubsection{Travel Companions}
% trip accompanies have siginificant effects on both location and trip rememberance. 
% location rememberance: travel alone -  100%, not alone - 55%, $\chi^2(1,94) = 31.812$, p < 0.01
% trip rememberance: travel alone - 62%, not alone - 90%, $\chi^2(1,133) = 9.016$, p < 0.01
Our study shows that both location and trip remembrances are influenced by the presence of travel companions, yet in a opposite way. For location rememberance, participants who believed that they traveled alone reported 100\% successful recalls of locations compare to 55\% rate reported by participants who believed that they traveled in groups, with $\chi^2(1,94) = 31.812$, $p < 0.00001$. On the other hand, trip details tend to be more easily remembered during accompanied visits as one participant recalled ``I remembered my friend and i walked a certain way -- I don't usually take this route". Participants who believed that they traveled alone were able to remember the exact trip 62\% of the time from watching Rewinds, yet those who believed that they traveled with companions could remember the exact trip 90\% of the time, with $\chi^2(1,133) = 9.016$, p = 0.0013 $(< 0.01)$.

@@

%TODO where to put this????  maybe split discussion into three sections on mem/believ/desir
% Jeff thinks that the discussion is already a bit long. It's fine in here, but maybe shorten it. Don't add more sections
% h - work on integrating this part or just deleting it
% \subsection{Episodic Memory}
% It could possibly be inferred that participants are likely to benefit more from Rewinds than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip rememberance when presented with static images as opposed to what they were asked after watchiing Rewinds, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip; since if participants fail to form any cognitive association with the locations depicted in the images, it is very unlikely that they would be able to associate target itineraries with the locations shown. Thus, it is perhaps safe to presume that location rememberance with images bounds above the maximum likelihood of trip rememberance with the same images presented to the participants, and we used location rememberance from image recognition as an upper-limit gauge against trip rememberance from Rewinds thereafter.

% One plausible explanation of the improvement may be inherent to the encoding of episodic memory in the first place. As suggested by the levels-of-processing theory(ref), encoding efficacy is heavily dependent on the levels to which stimuli are being actively or passively processed. Not surprisingly, superficial mental engagement would produce weaker, and less enduring cognitive representation which would demand more contextually accurate and abundant cues for memory triggery, and vice versa. Oftentimes, the cognitive representation resulted from shallow processing may not even be retrievable at all as it is simply not available for retrieval.  As previously described, both of the trips revealed in our results, those that were taken farthest away from home and those that were least frequently taken, may not be special enough to engage deep memory encoding nor often enough to reinforce incidental remembrance. This means that it would be more difficult for participants to recall these trips if the provided cues are not effective enough. Since Rewinds provide richer and more contextually meaningful cues than static street-side images, it is reasonable that Rewinds significantly outperforms plain images as suggested by our statistical analysis. While being asked about location and trip rememberance on one of his least frequently taken routes, P7 was not able to identify the location from the static image at first; yet he recognized the trip from Rewinds as ``one street (in the video) helps me (him) remember". 
% The Effects of Divided Attention on Encoding and Retrieval Processes in Human Memory 
% http://www2.psych.ubc.ca/~pgraf/Psy583Readings/Craik%20et%20al%201996.pdf
% Craik et al. 1996

% The extent to which attention is dedicated while taking a trip may be another possible explanation for the observed improvement using Rewinds. This is perhaps particularly applicable to explain why participants who were self-identified regular GPS users tend to benefit more from Rewinds with episodic memory recalling. Laboratory work done by Craik et al. 1996 (ref) has demonstrated that the encoding of memory is more effective under full-attention conditions than divided-attention conditions.  This means that regular GPS users, who are often directed step-by-step on how to navigate through their environments, may be less likely to engage their attention while taking their routes, leading to less available cognitive representation for retrieval. P3, a self-identified regular GPS user who had 30\% recalling rate using static images and 80\% trip remembrance rate using Rewinds said ``I really don't pay attention to my surroundings when I'm driving. (I am) Focused (focused) on GPS and road"  Under this circumstance, it is almost certain that Rewinds would provide a more successful access to episodic memory as it provides more spatial-temporal cues for memory retrieval, not only by adjusting image attributes to reflect timely and weather changes, but also by mimicing the motions of the users\' when they travel down a particular path.

% h - I'll comment it back if it is necessary
% It may appear to be puzzling that participants seem to recall less about specifics of their trips from Rewinds compared to static imageries when they traveled alone, with a significant decrease around 40\%, $\chi^2(1, N=168) = 33.831$, $p <.01$; when the trips were most frequently taken, with a significant decrease of 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$; and when the trips taken were 10 - 100 meters (33 - 328 feet) away from home, with a mildly significant decrease of 50\%, $\chi^2(1, N=16) = 2.25$, $p <.1$. One possible reason for this would be it is hard for participants to tie a repeatedly taken trip with a specific instance in time, and trips taken near home could also be the ones that are most frequently taken. However, since our study was using location rememberance from images as an upper-limit estimation about participants' trip rememberance from the images, it is also likely that the actual trip rememberance from images could be lower.
